\section{Formal results}
\label{sec:soundness}

In this section, we show that our flow-sensitive enforcement can be
embedded into the flow-insensitive version of LIO. Additionally, we
provide security guarantees in terms of non-interference definitions by reusing
previous results about LIO.

\subsection{Embedding into \lio{}}

We now show that we can implement our flow-sensitive semantics in terms of
flow-insensitive references. More concretely, every flow-sensitive reference
with label |l_d| created in a context where the current label is |l_o| (and
thus stored in |mS| as |LabeledTCB l_o (LabeledTCB l_d t)|), can be represented
by a flow-insensitive reference with label |l_o|, whose contents are
another flow-insensitive reference containing |t| and labeled |l_d|.

Figure~\ref{fig:fs-exts-semantics-impl} shows our implementation of
the flow-sensitive reference operations in this setting.  For a given
store |c|, we define the |impl (-) (c)| function, which given a term
|t| in \liofs{}, produces a term |impl (t) (c)| in \lio{}, which is
obtained by expanding the definitions of flow-sensitive operations in
terms of flow-insensitive ones. This function is applied
homomorphically in all other cases. We use the |WrapRef| to mark
the flow-insensitive references that are being used to represent
flow-sensitive ones, so as to distinguish them from ordinary
flow-insensitive references. The functions |wrap| and |unwrap| are
used to add and remove this boundary encoding.
%
In the embedding of |writeRef S|, we use |toLabeled| to make any
changes to the current label (possibly caused by reading the outer
reference) local to this operation. Inside |toLabeled|, the code
fetches the inner reference (|readRef I|), and then performs the
actual write of the new value.  If this fails, the computation
diverges, but, importantly, the current label was raised (with
|readRef I|) to reflect the fact that the label on the label of the
reference was observed.
%
The embedding of |upgradeR| follows similarly, but further relies on
|upgradeRI| which, like the upgrade function for labeled values,
creates a \emph{new} reference with the same contents as the supplied
reference |i|, but whose label is upgraded.
%
For the sake of brevity, we do not explain the mapping in further
detail, but remark that it is directly based on the rules in
Figure~\ref{fig:sos:fs}.
%

\begin{figure}
\small
\begin{code}
wrap r @= WrapRef r
unwrap (WrapRef r) @= r

impl (LIORefTCB S r) ((lcurr,mI,mS)) @= wrap (LIORefTCB I (labelOfR I mS(r)) r)

impl (newRef S) c @= \ l t . do
  i      <- newRef I l t
  lcurr  <- getLabel
  o      <- newRef I lcurr i
  return (wrap o)


impl (readRef S) c @= \r . readRef I (unwrap r) >>= readRef I


impl (writeRef S) c @= \r t .let o = unwrap r in do
  lcurr <- getLabel
  toLabeled (lcurr lub (labelOf o)) $ do
    i <- readRef I o
    writeRef I i t


impl (labelOfR S) c @= \r . 
  readRef I (unwrap r) >>= return . labelOfR I

impl (upgradeR) c @= \r l . let  o   = unwrap r
                                 l'  = labelOf o in do
  lcurr <- getLabel
  if (lcurr cantFlowTo l) then diverge else 
    toLabeled l' $ do
      i   <- readRef I o
      i'  <- upgradeRI i l
      writeRef I o i'

impl (withRefs v t) ((lcurr,mI,mS)) @= impl t ((lcurr, mI, mS'))
   where
    mS'  = {a mapsto mS(a) inlinesep a element dom mS and (addrs v) }

\end{code}
\caption{Implementation mapping for flow-sensitive references. For all other terms, the function is applied homomorphically.\label{fig:fs-exts-semantics-impl}}
\end{figure}

We extend this definition naturally to convert \liofs{} environments
into \lio{} environments, by having |impl1 ((lcurr, mI, mS)) @=
(lcurr, mI')| where |mI' = mI[ a1 mapsto v1, ..., a_n mapsto v_n ]|
for each binding of the form |a_i mapsto LabeledTCB li v_i| in
|mS|. Note that the domains of |mI| and |mS| are disjoint because the
|fresh| predicate that we use in the semantics is assumed to produce
globally unique addresses.

In order to prove that our implementation is correct with respect to
the semantics, we show that, if we take a program with flow-sensitive
operations, and expand those operations, replacing them by the code in
Figure~\ref{fig:fs-exts-semantics-impl}, then its behavior corresponds
with the flow-sensitive semantics.

% For the proof, we will require a lemma stating that, for every
% flow-sensitive reference, the label-on-the-label always flows to the
% label of the reference at every program point.

% \begin{lemma} Let |t| be a term in \liofs{}, |c = (lcurr, mI, mS)| an
%   initial environment and |r| a flow-sensitive reference in |c|.
%   Then, for every configuration |conf ((lcurr',mI',mS')) t'| reachable
%   from |conf c t|, if we consider |mS'(r) = LabeledTCB l_o (LabeledTCB
%   l_d v)|, we have that |l_o canFlowTo l_d|.
% \end{lemma}

% We can now state our equivalence theorem.

\begin{theorem}[Embedding \liofs{} in \lio{}]
\label{thm:eq} Let |t| be a well-typed term in \liofs{}.
  Then if |conf c t ==> conf c' v|, we have |conf (impl1 c) (impl t c)
  ==> conf (impl1 c') (impl v c)|, and if 
  |conf c t ==> conf c' diverge|, then
  |conf (impl1 c) (impl t c) ==> conf (impl1 c') diverge|.
  % \textrm{\hl{this is not right:
  %     the constructors for FS references are different (one wraps FI
  %     references, the other doesn't), so syntactic equiality is not
  %     right.  we need the equivalence relation I defined before}}
\end{theorem}
%
While straight forward, this theorem highlights an important result: in
floating label systems, flow-sensitive references can be encoded in a calculus
with flow-insensitive references and explicitly labeled values.

\subsection{Security guarantees for \liofs{}, \lioafs{} and \lioconc{}}

From previous results, we know that LIO satisfies termination-insensitive
non-interference (TINI) in the sequential setting, and termination-sensitive
non-interference (TSNI) in the concurrent setting. By using the
embedding theorem we can extend these results for LIO with flow-sensitive references.

For completeness, we now present our
non-interference theorems, as straightforward applications of the theorems in
previous work. Our security results rely on the notion of
|l|-equivalence for terms and configurations, which captures the idea
of terms that cannot be distinguished by an attacker which can observe
data at level |l|. A pair of terms |t1, t2| is said to be
|l|-equivalent (written |t1 &~ t2|) if, after erasing all the
information more sensitive than |l| from |t1| and |t2|, we obtain
syntactically equivalent terms. This definition extends naturally to
configurations.

Intuitively, non-interference means that an attacker at level |l|
cannot distinguish among different runs of a program with
|l|-equivalent initial configurations.

\begin{theorem}[TINI for \liofs{}]
  Consider two well-typed terms |t1| and |t2| in
  \liofs{} which do not contain any |cdot TCB| syntax nodes, such that |t1 &~
  t2|, where |l| is the attacker observation level.
  Let |c| be an initial environment, and let
\[
    |conf c (t1) ==> conf c1 v1|\mbox{ and }
    |conf c (t2) ==> conf c2 v2|
\]
  Then, we have that |conf c1 v1 &~ conf c2 v2|.
\end{theorem}

\begin{proof}
  By expanding all the flow-sensitive operations in |t1| and |t2| using their
  definition given in Figure~\ref{fig:fs-exts-semantics-impl}, we get
  terms in \lio{}, which by Theorem~\ref{thm:eq} has equivalent
  semantics. Therefore, the result follows from the
  \lio{} TINI result of~\citep{stefan:lio}.
\end{proof}

\begin{corollary}[TINI for \lioafs{}]
  The previous non-interference result can be easily extended to \lioafs{}. 
  In \lioafs{}, the |unlabel| operation
  triggers the automatic upgrades mechanism, which 
  performs the |upgrade| operation for every flow-sensitive
  reference in scope before actually raising the current label.
  Regardless of how |unlabel| is used, we note that the resulting term
  (after inserting the necessary |upgrade|s), is just an \liofs{}
  term. Therefore, the main TINI result for \liofs{} applies.

 % Additionally, the |withRefs| primitive allows the programmer
 %  to restrict the set of references in scope, reducing the number of
 %  |upgrade|s that need to be performed when |unlabel| occurs.

% \hl{move
%    withRefs}
\end{corollary}

For the concurrent result, we need a supporting lemma which states
that the current label is always at least as sensitive as
the label of every reference in scope. Formally,

\begin{lemma} \label{lem:lolFlowslCurr}
  Let |t| be a well-typed term in \lioconc{}, |c = (lcurr, mI, mS)| an initial
  environment and |a| the address of a flow-sensitive reference |r| in |c|,
  where |mS(a) = LabeledTCB l_o (LabeledTCB l_d v)|. Then, if |conf c t ==>
  conf (lcurr',mI',mS') t'|, we have that |l_o canFlowTo lcurr'|.
\end{lemma}

\begin{proof}
  Note that the result holds immediately after creating |r|, since
  the current label is the label on the label of |r|, i.e., |l_o =
  lcurr|. It is easy to show that |l_o| is immutable, since there are
  no reduction rules that modify it. Moreover, given that the current
  label is monotonic, the only way in which |l_o canFlowTo lcurr| can
  cease to hold is if |r| is accessed from a different thread. But in
  order to pass |r| to a different thread, a labeled object must be
  used as intermediary, and the label of such object would have to be
  at least |lcurr|, the current label in the thread that created
  |r|. As a result, if we were to pass |r| to another thread in this
  way, then the target thread would also have to be tainted by
  |lcurr|, and the result would still hold.
\end{proof}

We can now prove our non-interference theorem for \lioconc. This result is
slightly stronger than TINI, since it implies that there can be no termination
or internal timing leaks.

\begin{theorem}[TSNI for \lioconc{}]
  Consider two well-typed terms |t1| and |t2| in
  \lioconc{} which do not contain any |cdot TCB| syntax nodes, such that |t1
  &~ t2|, where |l| is the attacker observation level.
  Let |c = (lcurr, mI, mS)| be an initial environment, and let
\[
 |tconf mI mS (thread lcurr (refs (dom mS)) t1) ==> M1|
\]
  Then, there exists some configuration |M2| such that |tconf mI mS (thread lcurr (refs (dom mS)) t2)
  ==> M2| and |M1 &~ M2|.
\end{theorem}

\begin{proof} Because of Lemma~\ref{lem:lolFlowslCurr}, and
  looking at the embedding of |writeRef S| and |upgradeR|, we note
  that the first |readRef I| operation in each |toLabeled| block will
  be trying to raise the current label to |l|. However, since |l
  canFlowTo lcurr|, these operations will never effectively raise the
  current label. This means that using |toLabeled| is not necessary to
  preserve the semantics, because there is no need to restore the
  current label afterwards. As a result, and after removing
  |toLabeled| in these two cases, we note that the embedding produces
  valid concurrent \lio{} terms (which does not have |toLabeled|).


%%   However, we note that all the critical sections in our operations consist in
%%   a |readRef I| followed by a |writeRef I|. Therefore, we argue that
%%   the problem could be addressed by replacing these operations with
%%   |atomicModifyRef I|, an atomic operation that performs the read and write and
%%   is part of the actual implementation of LIO, but which we elide in order to
%%   simplify our presentation.

  Finally, by expanding all the flow-sensitive operations in |t1| and |t2|
  using their definition given in
  Figure~\ref{fig:fs-exts-semantics-impl}, we get terms in concurrent
  \lio{}.  Therefore, the result follows from the
  termination-sensitive non-interference of concurrent
  \lio{}~\citep{stefan:addressing-covert}.
  %
  We remark, however, that our embedding includes no synchronization to
  ensure atomicity of the flow-sensitive operations, so certain
  interleavings that break semantic equivalence are possible.
  Importantly, this does not affect security.
\end{proof}

The detailed proofs for the results in this section can be found in the
extended version of the paper~\cite{extended}.

\subsection{Permissiveness}
In Section~\ref{sec:related} we compare the permissiveness of our
system with previous flow-sensitive IFC systems. Here, we
solely remark that the above results imply that our
flow-sensitive calculus is as permissive as flow-insensitive
LIO. In particular, any flow-insensitive LIO program can be trivially
converted to a flow-sensitive LIO program (without auto-upgrades) by
using flow-sensitive references instead of flow-insensitive ones.
Since these references would never be upgraded, they will behave just
like their flow-insensitive counterparts.
%
This means that all existing LIO programs can be run in our
flow-sensitive monitor. This includes Hails~\cite{giffin:hails}, a
web framework using LIO, on top of which a number of applications
have been built (e.g., GitStar\footnote{\url{www.gitstar.com}}, a
code-hosting web platform,
LearnByHacking\footnote{\url{www.learnbyhacking.org}}, a
blog/tutorial platform similar to School of Haskell, and
LambdaChair~\cite{stefan:2012:arxiv-flexible}, an EasyChair-like
conference review system).


% GitStar's core could be run in our
%   flow-sensitive monitor; the full system uses other features like
%   privileges and exceptions that we have not directly implemented in
%   this version of the flow-sensitive monitor, but we see no major
%   challenges in incorporating flow-sensitivity to the full version of
%   LIO.

% \hl{NOTE}

% Why is |toLabeled| secure in the implementation of |writeRef S|? It
% would seem that it is impossible for the first |readRef I| to
% effectively raise the current label, i.e. if a reference with LOL |l|
% is in scope, it is always the case that |l canFlowTo lcurr|. Why? If a
% reference has LOL |l|, it must have been created in a context where
% the current label was |l|. From here, it is not possible to access the
% reference from a context with a lower current label. This would
% suggest that |toLabeled| is not necessary, since |readRef| is not
% changing the current label anyway.

% Local Variables:
% TeX-master: "main.lhs.tex"
% TeX-command-default: "Make"
% End:
